---
title: "Foundations: Vectors"
output: rmarkdown::github_document
---

# Quiz

1. What are the four common types of atomic vectors? What are the two rare types?

> logical, integer, double, and character (two rarer types are complex and raw)

2. What are attributes? How do you get them and set them?

> let you add additional metadata to an object; set them with attr() or attributes()

3. How is a list different from an atomic vector? How is a matrix different from a data frame?

> All the elements of an atomic vector must be the same type, not so for lists. Similarly, all elements of a matrix must be the same type but a data frame can contain multiple types in different columns.

4. Can you have a list that is a matrix? Can a data frame have a column that is a matrix?

> Copied from the answers at the end of the chapter: "You can make a “list-array” by assigning dimensions to a list. You can make a matrix a column of a data frame with df$x <- matrix(), or by using I() when creating a new data frame data.frame(x = I(matrix()))."

5. How do tibbles behave differently from data frames?

> better printing, stringsAsFactors = FALSE, drop = FALSE as defaults; they don't transform non-syntactic names; they don't coerce input types; tibbles won't recycle values for other columns unless they are of length 1; tibbles allow you to build on columns as they are constructed; tibbles do not support row names on purpose; tibbles don't do partial matching with $ column selecting; easier creation of list columns.

## 3.2 Atomic vectors

1. How do you create raw and complex scalars? (See ?raw and ?complex)

> Use as.raw/as.complex

2. Test your knowledge of the vector coercion rules by predicting the output of the following uses of 

```{r}
c(1, FALSE) # 1, 0
c("a", 1) # "a", "1"
c(TRUE, 1L) # 1 1 
``` 

3. Why is 1 == "1" true? Why is -1 < FALSE true? Why is "one" < 2 false?

> coercion

4. Why is the default missing value, NA, a logical vector? What’s special about logical vectors? (Hint: think about c(FALSE, NA_character_).)

> Logical vectors take precedence over all ather types. All other vector types can be coerced to logical/from binaries

5. Precisely what do is.atomic(), is.numeric(), and is.vector() test for?
 
> is.atomic - tests whether an object is atomic (may not be a vector), is.numeric - tests for modes and implicit classes, not type, and  is.vector - return false for any mode that has attributes other than names

## 3.3 Attributes

1. How is setNames() implemented? How is unname() implemented? Read the source code.

> Ha: `names(object) <- nm; object`

2. What does dim() return when applied to a 1D vector? When might you use NROW() or NCOL()?

> NULL; 
> "nrow and ncol return the number of rows or columns present in x. NCOL and NROW do the same treating a vector as 1-column matrix"

```{r}
x <- c(1, 2, 3)
dim(x)
NROW(x)
NCOL(x)
```

3. How would you describe the following three objects? What makes them different from 1:5?

```{r}
x1 <- array(1:5, c(1, 1, 5))
x1
```

```{r}
x2 <- array(1:5, c(1, 5, 1))
x2
```

```{r}
x3 <- array(1:5, c(5, 1, 1))
x3
```

> FIXME: not sure what the best way to describe these is... They put the values 1:5 into the 3 possible dimensions.
> FIXME: 1:5 doesn't technically have dimensions and is not an array

4. An early draft used this code to illustrate structure():

```{r}
structure(1:5, comment = "my attribute")
```

But when you print that object you don’t see the comment attribute. Why? Is the attribute missing, or is there something else special about it? (Hint: try using help.)

> from ?attr: "Note that some attributes (namely class, comment, dim, dimnames, names, row.names and tsp) are treated specially and have restrictions on the values which can be set."

## 3.4 S3 Atomic vectors

1. What sort of object does table() return? What is its type? What attributes does it have? How does the dimensionality change as you tabulate more variables?

```{r}
x <- table(factor(c("a", "b")))
x
str(x)
attributes(x)

x <- table(factor(c("a", "b", "c")))
x
attributes(x)
```

2. What happens to a factor when you modify its levels?

```{r}
f1 <- factor(letters)
f1
str(f1)
attributes(f1)
as.numeric(f1)

levels(f1) <- rev(levels(f1))
f1
str(f1)
attributes(f1)
as.numeric(f1)
```

> The ordering of the `levels` attribute changes but the underlying values do not.

3. What does this code do? How do f2 and f3 differ from f1?

```{r}
f2 <- rev(factor(letters))
f2
as.numeric(f2)
```

> f2 has reversed underlying values an levels in a different order

```{r}
f3 <- factor(letters, levels = rev(letters))
f3
as.numeric(f1)
as.numeric(f3)
levels(f1)
levels(f3)
```

> f3 has reversed underlying values but levels in the same order

## 3.5 Lists

1. List all the ways that a list differs from an atomic vector.

> FIXME: Lists can contain references to objects of multiple types; lists always contain references to objects; lists can be recursive and contain other lists...

2. Why do you need to use unlist() to convert a list to an atomic vector? Why doesn’t as.vector() work?

> FIXME: the rules for coercion are complex and require their own function

3. Compare and contrast c() and unlist() when combining a date and date-time into a single vector.

```{r}
x <- as.POSIXct("2018-08-01 22:00")
y <- as.Date("1jan2018", "%d%b%Y")
z <- list(x, y)
c(z)
unlist(z)
```

> c() keeps a list where each element refers to an object of a different class
> unlist() coerces the elements to numeric

## 3.6 Data Frames and tibbles

1. Can you have a data frame with zero rows? What about zero columns?

> rows: yes, columns: no

2. What happens if you attempt to set rownames that are not unique?

```{r}
x <- data.frame(1:10)
# rownames(x) <- c(1, 1:9)
# Error in `.rowNamesDF<-`(x, value = value) :
#   duplicate 'row.names' are not allowed
```

3. If df is a data frame, what can you say about t(df), and t(t(df))? Perform some experiments, making sure to try different column types.

```{r}
df <- data.frame(x = 1:10, y = 1:10)
t(df)
t(t(df))
```

> The type is coerced

 4. What does as.matrix() do when applied to a data frame with columns of different types? How does it differ from data.matrix()?

```{r}
df <- data.frame(x = 1:10, y = letters[1:10])
as.matrix(df)
data.matrix(df)
```

> Different coercion rules. From the help: data.matrix: "Return the matrix obtained by converting all the variables in a data frame to numeric mode and then binding them together as the columns of a matrix. Factors and ordered factors are replaced by their internal codes."
