---
title: "Foundations: Subsetting"
output: rmarkdown::github_document
---

## 4.2 Selecting multiple elements

1. Fix each of the following common data frame subsetting errors:

```{r, eval=FALSE}
mtcars[mtcars$cyl = 4, ]
mtcars[-1:4, ]
mtcars[mtcars$cyl <= 5]
mtcars[mtcars$cyl == 4 | 6, ]
```

```{r}
mtcars[mtcars$cyl == 4, ]
mtcars[-c(1:4), ]
mtcars[mtcars$cyl <= 5, ]
mtcars[mtcars$cyl %in% c(4, 6), ]
# or
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
```

2. Why does the following code yield five missing values? (Hint: why is it different from x[NA_real_]?)

```{r}
x <- 1:5
x[NA]
x[NA_real_]
```

> FIXME

3. What does upper.tri() return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?

```{r}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```

> It returns the upper triangle of the matrix.
> Second part (FIXME?): it behaves like drop = TRUE?

4. Why does mtcars[1:20] return an error? How does it differ from the similar mtcars[1:20, ]?

> Indexing a data frame like that behaves like a list so it tries to select the first twenty columns of the data frame but there are not that many columns. The second option correctly selects the first twenty rows.

5.Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).

> 

6. What does df[is.na(df)] <- 0 do? How does it work?

> Replaces all NAs with 0s


## 4.3 Selecting a single element

1. Brainstorm as many ways as possible to extract the third value from the cyl variable in the mtcars dataset.

> 

2. Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Then extract the R squared from the model summary (summary(mod)).

> 

## 4.5 Applications

1. Positive integers select elements at specific positions, negative integers drop elements; logical vectors keep elements at positions corresponding to TRUE; character vectors select elements with matching names.

> 

2. [ selects sub-lists: it always returns a list. If you use it with a single positive integer, it returns a list of length one. [[ selects an element within a list. $ is a convenient shorthand: x$y is equivalent to x[["y"]].

> 

3. Use drop = FALSE if you are subsetting a matrix, array, or data frame and you want to preserve the original dimensions. You should almost always use it when subsetting inside a function.

> 

4. If x is a matrix, x[] <- 0 will replace every element with 0, keeping the same number of rows and columns. In contrast, x <- 0 completely replaces the matrix with the value 0.

> 

5. A named character vector can act as a simple lookup table: c(x = 1, y = 2, z = 3)[c("y", "z", "x")]

> 
