---
title: "Foundations: Control Flow"
output: rmarkdown::github_document
---

## 5.2 Choices

1. What type of vector does each of the following calls to `ifelse()` return?

```{r}
x <- ifelse(TRUE, 1, "no")
typeof(x)
```
> The `1` is returned, and interpreted as a `double`. To make it an integer, you would have to use `1L`.

```{r}
y <- ifelse(FALSE, 1, "no")
typeof(y)
```
> The `"no"` is returned, which is of type `character`.

```{r}
z <- ifelse(NA, 1, "no")
typeof(z)
```
> `NA` is returned, which is of type `logical`. Interestingly, you can do the following and get the same result:

```{r}
z <- ifelse(NA)
typeof(z)
```

2. Why does the following code work?

```{r}
x <- 1:10
if (length(x)) "not empty" else "empty"

x <- numeric()
if (length(x)) "not empty" else "empty"
```
> The expression `length(x)` evaluates to an `integer` which is coerced to a `logical`. Zero is coerced to `FALSE` and any other value to `TRUE`. This integer-to-logical coercion is a feature of C that was carried over to R.

## 5.3 Loops

1. Why does this code succeed without errors or warnings?

```{r}
x <- numeric()
out <- vector("list", length(x))
for (i in 1:length(x)) {
  out[i] <- x[i] ^ 2
}
out
```
> The expression `1:length(x)` evaluates to `1:0` in this case, which is allowed in R. The code inside the `for` loop will evaluate `out[1] <- NA` when `i == 1` and when `i == 0` it will evaluate `numeric(0) <- numeric(0)` which is meaningless but doesn't produce an error. Use `seq_along` instead:

```{r}
x <- numeric()
out <- vector("list", length(x))
for (i in seq_along(x)) {
  out[i] <- x[i] ^ 2
}
out
```

2. When the following code is evaluated, what can you say about the vector being iterated?

```{r}
xs <- c(1, 2, 3)
for (x in xs) {
  xs <- c(xs, x * 2)
}
xs
```
> The vector is being appended to, one new element per iteration.

3. What does the following code tell you about when the index is updated?

```{r}
for (i in 1:3) {
  i <- i * 2
  print(i) 
}
```
> `i` is updated (reset) at the beginning of each iteration to the value given in the parentheses. Any modifications to `i` inside the loop are forgotten at the end of the iteration.


