---
title: "Foundations: Functions"
output: rmarkdown::github_document
---

## 6.2 Function fundamentals

1. Given a name, like `"mean"`, `match.fun()` lets you find a function. Given a function, can you find its name? Why doesn’t that make sense in R?

> R's variables and functions are references to objects, so a function can have more than one reference pointing to it and therefore there is no unique name:

```{r}
x <- y <- z <- function(x){
  x^2
}
```

2. It's possible (although typically not useful) to call an anonymous function. Which of the two approaches below is correct? Why?

```{r error = TRUE}
function(x) 3()
(function(x) 3)()
```
> The first example is trying to use the number `3` as a function name which is not allowed in R The second (correct) example creates an anonymous function `function(x) 3` and calls it using the two sets of parentheses.

3. A good rule of thumb is that an anonymous function should fit on one line and shouldn’t need to use `{}`. Review your code. Where could you have used an anonymous function instead of a named function? Where should you have used a named function instead of an anonymous function?

> In general, anaoymous functions would be used inside other function calls where a (simple) function is applied to elements of a vector or list. Examples would be inside the R-base `apply` functions and the `purrr::map` functions. e.g.:

```{r}
j <- list(1, 2, 3)
lapply(j, function(x) x^2)
purrr::map(j, function(x) {y <- x^2;y*2})
```

4. What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?

```{r}
a <- function(x){x^2}
is.function(a)
is.primitive(a)
is.primitive(`+`)
```

5. This code makes a list of all functions in the base package:

```{r}
objs <- mget(ls("package:base", all = TRUE), inherits = TRUE)
funs <- Filter(is.function, objs)
```

> The functions below were used to create output to answer questions that follow:

```{r message = FALSE, warning = FALSE}
library(tibble)
library(dplyr)
library(purrr)

gn <- function(lst1){
  # Extract only NULL elements' names from list lst1
  j <- list()
  ind <- 0
  for(i in seq_along(lst1)){
    if(is.null(lst1[[i]])){
      j[[ind <- ind + 1]] <- names(lst1)[i]
    }
  }
  j
}

package.funs <- function(p = "base"){
  objs <- mget(ls(paste0("package:", p)), inherits = TRUE)
  funs <- Filter(is.function, objs)
  suppressWarnings(
    all.funcs <- purrr::map(funs, function(x){formals(args(x))})
  )
  # special.funcs is a vector of names of functions with NULL formals
  special.funcs <- gn(all.funcs) %>%
    map_chr(1)
  special.funcs <- as.data.frame(special.funcs) %>% 
    as_tibble() %>% 
    rename(name = special.funcs) %>%
    mutate(name = as.character(name),
           num.args = NA,
           is.primitive = sapply(name, function(x) is.primitive(get(x))))
  args.funcs <- all.funcs
  # Remove NULL entries (they are already in special.funcs)
  args.funcs[sapply(args.funcs, is.null)] <- NULL
  # Now, args.funcs contains only functions that have formals reported
  # Extract the number of arguments for each function
  num.args <- purrr::map(args.funcs, length) %>%
    map_int(1)
  args.funcs <- as.data.frame(num.args) %>%
    as_tibble(rownames = "name") %>%
    # Can't use mutate(is.primitive = is.primitive(get(.))) here
    # Check out https://github.com/tidyverse/dplyr/issues/4280 to see why
    mutate(is.primitive = map_lgl(name,~is.primitive(get(.)))) %>%
    arrange(desc(num.args))
  bind_rows(args.funcs, special.funcs)
}
a <- package.funs()
```

> There are a couple things to note here. The `formals` function does not return the arguments for primitives, but `NULL` instead. To get around this, the code above uses `formals(args())` as this will return arguments for both primitives and non-primitives. There are some builtin functions which you cannot get argument lists for even with `formals(args())`. These are returned in a vector called `special.funcs` from the function `package.funcs()` above.

  a. Which base function has the most arguments?

  > The first one in the `a` dataframe, `r a$name[1]`(), with `r a$num.args[1]` arguments.

  b. How many base functions have no arguments? What’s special about those functions?

  > It is (maybe) unknown how many base functions have zero arguments. There are `r length(which(is.na(a$num.args)))` functions which do not report their arguments using `formals()` or `formals(args())`.

  c. How could you adapt the code to find all primitive functions?

  > See the code above. It adds a column for this using these two lines:
  `is.primitive = sapply(name, function(x) is.primitive(get(x)))` and
  `mutate(is.primitive = map_lgl(name,~is.primitive(get(.)))) %>%`

6. What are the three important components of a function?

> The `formals`, `body`, and `environment`.

7. When does printing a function not show the environment it was created in?

> Primitive functions have `NULL` values for `formals`, `body`, and `environment` attributes.

## 6.4 Lexical Scoping

1. What does the following code return? Why? Describe how each of the three c’s is interpreted.

```{r}
c <- 10
c(c = c)
```

> The first `c` is a variable and assigned a value, the `c()` is the vector combine function, the `c = ` on the second line is another variable inside the scope of the `c()` call, the `c` it is assigned to is the `c` from the first line, because it is up one level in scope.

2. What are the four principles that govern how R looks for values?

> 1. Name masking
  1. Functions vs. variables
  1. A fresh start
  1. Dynamic lookup

3. What does the following function return? Make a prediction before running the code yourself.

```{r}
f <- function(x) {
  f <- function(x) {
    f <- function() {
      x ^ 2
    }
    f() + 1
  }
  f(x) * 2
}
f(10)
```
> I predicted 202

## 6.5 Lazy Evaluation

1. What important property of `&&` makes `x_ok()` work?

```{r}
x_ok <- function(x) {
  !is.null(x) && length(x) == 1 && x > 0
}
x_ok(NULL)
x_ok(1)
x_ok(1:3)
```

>  Lazy evaluation. If x is `NULL`, only the `is.null(x)` part of the expression is evaluated.

What is different with this code? Why is this behaviour undesirable here?

```{r}
x_ok <- function(x) {
  !is.null(x) & length(x) == 1 & x > 0
}
x_ok(NULL)
x_ok(1)
x_ok(1:3)
```

>  The operator `&` is vectorized and therefore does not incorporate lazy evaluation. All parts of the expression will be evaluated.

2. What does this function return? Why? Which principle does it illustrate?

```{r}
f2 <- function(x = z) {
  z <- 100
  x
}
f2()
```

> It will return 100, because the argument list is evaluated lazily and `x` is set to 100 by using the value of `z` inside the function body.

3. What does this function return? Why? Which principle does it illustrate?

```{r}
y <- 10
f1 <- function(x = {y <- 1; 2}, y = 0) {
  c(x, y)
}
f1()
y
```

> In this case, `y` is set to 1 and `x` to 2 inside the argument list. Outside the function, `y` remains as 10. This is due to scoping rules.

4. In hist(), the default value of xlim is range(breaks), the default value for breaks is "Sturges", and

```{r}
range("Sturges")
```

Explain how hist() works to get a correct xlim value.

> `breaks` is recalculated inside the function and lazy evaluation applies the newly calculated values to the `xlim` argument.

5. Explain why this function works. Why is it confusing?

```{r}
show_time <- function(x = stop("Error!")) {
  stop <- function(...) Sys.time()
  print(x)
}
show_time()
```

> Lazy evaluation binds x to be the new value of the stop function as defined inside the function body. This should not be written this way because it is changing the function definition for the base `stop` function to something that doesn't resemble stopping. Note this only happens inside the function's scope, so the `stop` function isn't changed globally. It relies on lazy evaluation to reduce the number of lines of code necessary so although clever, it is horrible practice.

## 6.6 Dot..dot..dot

1. Explain the following results:

```{r}
sum(1, 2, 3)
mean(1, 2, 3)
sum(1, 2, 3, na.omit = TRUE)
mean(1, 2, 3, na.omit = TRUE)
```

> `sum` is implemented using `...` as its first argument, so all values are used in the sum and it is correct.
`mean` is implemented using `x` as its first argument and therefore cannot do the mean of numbers unless they are in a vector. `na.omit` is not an argument for either of these functions. It is just passed along in `mean` because it is part of the `...` arguments but in `sum` it is coerced to `logical` and then to the type that is being summed. In this case it is coerced to `TRUE` then to `1` so the call becomes the equivalent of `sum(1, 2, 3, 1)` which evaluates to 7.

2. In the following call, explain how to find the documentation for the named arguments in the following function call:

```{r}
plot(1:10, col = "red", pch = 20, xlab = "x", col.lab = "blue")
```

> Looking at the help for `plot`, is says that the `...` parameters are explained in `par`.

3. Why does `plot(1:10, col = "red")` only colour the points, not the axes or labels? Read the source code of `plot.default()` to find out.

> `localAxis`, `localBox`, `localWindow`, and `localTitle` inside the `plot.default` function use the global `col` value which is the one set up in `par()`. The `col` passed into the function is used only to plot the points.

## 6.7 Exiting a function

1. What does `load()` return? Why don't you normally see these values?

> It returns the objects stored in the file that was loaded. These could be huge objects so you wouldn't want them printed to the screen.

2. What does `write.table()` return? What would be more useful?

> It returns`NULL`. A value representing success or failure of writing the file would be more useful.

3. How does the `chdir` parameter of `source()` compare to `in_dir()`? Why might you prefer one to the other?

> `in_dir()` takes code as an argument, leaving no question as to what is executed in the new directory. `source()` uses `chdir` to change directories somewhere in the function and it is difficult to trace how it works. It is better for code encapsulation purposes to use `in_dir()`.

4. Write a function that opens a graphics device, runs the supplied code, and closes the graphics device (always, regardless of whether or not the plotting code works).

```{r}
pl <- function(){
  plot.new()
  on.exit(dev.off(), add = TRUE)

  ## Plotting code here
}
```

5. We can use `on.exit()` to implement a simple version of `capture.output()`.

```{r}
capture.output2 <- function(code) {
  temp <- tempfile()
  on.exit(file.remove(temp), add = TRUE, after = TRUE)

  sink(temp)
  on.exit(sink(), add = TRUE, after = TRUE)

  force(code)
  readLines(temp)
}
capture.output2(cat("a", "b", "c", sep = "\n"))
```
Compare `capture.output()` to `capture.output2()`. How do the functions differ? What features have I removed to make the key ideas easier to see? How have I rewritten the key ideas so they’re easier to understand?

> The main difference is that `capture.output2()` uses multiple `on.exit` calls which make it much easier to see what is going on. These multiple calls are done in order and therefore the function flows in the natural order in which you would expect.

## 6.8 Function Forms

1. Rewrite the following code snippets into prefix form:

```{r}
1 + 2 + 3
`+`(3, `+`(1, 2))

1 + (2 + 3)
`+`(1, `+`(2, 3))

x <- 1:5
if (length(x) <= 5) x[[5]] else x[[n]]
`if`(length(x) <= 5, `[[`(x, 5), `[[`(x, n))
```

2. Clarify the following list of odd function calls:

```{r}
x <- sample(replace = TRUE, 20, x = c(1:10, NA))
```

> Here, the 20 is the `size` argument so this produces a vector of length 20 with replacement drawing from the vector given by `x`.

```{r}
y <- runif(min = 0, max = 1, 20)
```

> Here, the 20 is the `x` argument so this produces a vector of length 20 of draws from the unif distribution between 0 and 1.

```{r}
cor(m = "k", y = y, u = "p", x = x)
```

> Here, all arguments are at least partially named so it's easy to see what's going on. It's using `x` and `y` from the previous code and partial matching for arguments `u = use` and `m = method`. For `use`, partial matching is also used for the value, so `p = pairwise.complete.obs`. For `method`, the partial value match is `k = kendall`.

3. Explain why the following code fails:

```{r error = TRUE}
`modify<-` <- function(x, position, value) {
  x[position] <- value
  x
}
modify(get("x"), 1) <- 10
```

> This is because the return value of `get("x")` has been not been assigned to a variable and is temporary. You must assign it before doing this.

```{r}
a <- get("x")
modify(a, 1) <- 10
a
```

4. Create a replacement function that modifies a random location in a vector.

```{r}
`repl.rand<-` <- function(x, value){
  x[sample(seq_along(x), 1)] <- value
  x
}

x <- 1:10
repl.rand(x) <- 99
x
repl.rand(x) <- 55
x
```

5. Write your own version of `+` that pastes its inputs together if they are character vectors but behaves as usual otherwise. In other words, make this code work:

```{r error = TRUE}
1 + 2
"a" + "b"
`+` <- function(e1, e2){
  if(typeof(e1) == "character" || typeof(e2) == "character"){
    paste0(e1, e2)
  }else{
    tmp <- .Primitive("+")
    tmp(e1, e2)
  }
}
1 + 2
"a" + "b"
```

6. Create a list of all the replacement functions found in the base package. Which ones are primitive functions? (Hint: use `apropos()`)

```{r}
library(tibble)
library(purrr)
library(dplyr)
repl.funcs <- apropos("<-") %>%
  enframe(name = NULL) %>% 
    rename(name = value) %>% 
      mutate(is.primitive = map_lgl(name, ~is.primitive(get(.))))

are.primitive <- repl.funcs[repl.funcs$is.primitive,]
are.primitive
```

> `r nrow(are.primitive)` of the `r nrow(repl.funcs)` base replacement functions are primitive.

7. What are valid names for user-created infix functions?

> User infix functions are anything surrounded by %%, and can't contain %. Examples are `% %`, `%/\%`, `%XX%`.

8. Create an infix xor() operator.

```{r}
`%xor%` <- function(e1, e2){
    (e1 && !e2) || (!e1 && e2)
}
TRUE %xor% FALSE
TRUE %xor% TRUE
FALSE %xor% FALSE
FALSE %xor% TRUE
```

9. Create infix versions of the set functions `intersect()`, `union()`, and `setdiff()`. You might call them `%n%`, `%u%`, and `%/%` to match conventions from mathematics.

```{r}
`%n%` <- function(x, y){
  base::intersect(x, y)
}

`%u%` <- function(x, y){
  base::union(x, y)
}

`%/%` <- function(x, y){
  base::setdiff(x, y)
}

(x <- c(sort(sample(1:20, 9)), NA))
(y <- c(sort(sample(3:23, 7)), NA))
x %n% y
x %u% y
x %/% y
```
