---
title: "Foundations: Environments"
output: rmarkdown::github_document
---

## 7.2

```{r}
library(rlang)
````

1. List three ways in which an environment differs from a list.

> 1. Every name must be unique.
  1. The names in an environment are not ordered.
  1. An environment has a parent.
  1. Environments are not copied when modified.
  
2. Create an environment as illustrated by this picture.

```{r}
e1 <- env()
e1$loop <- e1
```

3. Create a pair of environments as illustrated by this picture.

```{r}
e1 <- env()
e2 <- env()
e1$loop <- e2
e2$dedoop <- e1
```

4. Explain why `e[[1]]` and `e[c("a", "b")]` don’t make sense when `e` is an environment.

> Environments are not ordered, so numbered or vectorized indexing make no sense.

5. Create a version of `env_poke()` that will only bind new names, never re-bind old names. Some programming languages only do this, and are known as single assignment languages.

```{r}
env_poke_single <- function(env = caller_env(), nm, value){
  stopifnot(is_string(nm))
  if(!env_has(env, nm)){
    assign(nm, value, envir = env)
  }
}
```

6. What does this function do? How does it differ from `<<-` and why might you prefer it?

```{r error = TRUE}
rebind <- function(name, value, env = caller_env()) {
  if(identical(env, empty_env())){
    stop("Can't find `", name, "`", call. = FALSE)
  }else if(env_has(env, name)){
    env_poke(env, name, value)
  }else{
    rebind(name, value, env_parent(env))
  }
}
rebind("a", 10)
a <- 5
rebind("a", 10)
a
```

> This is a recursive function that attempts to re-bind the value to the variable in the current environment, followed by the parent, followed by it's parent, etc. It re-binds in the first environment it finds where the variable exists. This will recurse all the way past the `.GlobalEnv` and through the package environments, unlike `<<-` which stops at the `.GlobalEnv` and creates the variable if it doesn't exist there.

## 7.3 Recursing over environments

1. Modify `where()` to return all environments that contain a binding for name. Carefully think through what type of object the function will need to return.

```{r}
where <- function(name, env = caller_env()) {
  if(identical(env, empty_env())) {
    NULL
  }else if(env_has(env, name)){
    c(env, where(name, env_parent(env)))
  }else{
    where(name, env_parent(env))
  }
}
```

2. Write a function called `fget()` that finds only function objects. It should have two arguments, `name` and `env`, and should obey the regular scoping rules for functions: if there’s an object with a matching name that’s not a function, look in the parent. For an added challenge, also add an `inherits` argument which controls whether the function recurses up the parents or only looks in one environment.

```{r}
fget <- function(name, env = caller_env(), inherits = FALSE){
  if(identical(env, empty_env())){
    NULL
  }else if(env_has(env, name) && is.function(env[[name]])){
    if(inherits){
      c(env, fget(name, env_parent(env), inherits))
    }else{
      env
    }
  }else{
    if(inherits){
      fget(name, env_parent(env), inherits)
    }
  }
}

# Make new environment e1 whose parent is R_GlobalEnv
e1 <-  env(a = 1, b = 2, c =3, mean = function(x)sum(x)/length(x))
# Make new environment e2 whose parent is e1
e2 <-  env(e1, a2 = 1, b2 = 2, c2 =3, mean = function(x)sum(x)/length(x))
fget("mean", e2, TRUE)
fget("mean", e2)
```
