---
title: "Foundations: Environments"
output: rmarkdown::github_document
---

## 7.2

```{r}
library(rlang)
````

1. List three ways in which an environment differs from a list.

> 1. Every name must be unique.
  1. The names in an environment are not ordered.
  1. An environment has a parent.
  1. Environments are not copied when modified.
  
2. Create an environment as illustrated by this picture.

```{r}
e1 <- env()
e1$loop <- e1
```

3. Create a pair of environments as illustrated by this picture.

```{r}
e1 <- env()
e2 <- env()
e1$loop <- e2
e2$dedoop <- e1
```

4. Explain why `e[[1]]` and `e[c("a", "b")]` don’t make sense when `e` is an environment.

> Environments are not ordered, so numbered or vectorized indexing make no sense.

5. Create a version of `env_poke()` that will only bind new names, never re-bind old names. Some programming languages only do this, and are known as single assignment languages.

```{r}
env_poke_single <- function(env = caller_env(), nm, value){
  stopifnot(is_string(nm))
  if(!env_has(env, nm)){
    assign(nm, value, envir = env)
  }
}
```

6. What does this function do? How does it differ from `<<-` and why might you prefer it?

```{r error = TRUE}
rebind <- function(name, value, env = caller_env()) {
  if(identical(env, empty_env())){
    stop("Can't find `", name, "`", call. = FALSE)
  }else if(env_has(env, name)){
    env_poke(env, name, value)
  }else{
    rebind(name, value, env_parent(env))
  }
}
rebind("a", 10)
a <- 5
rebind("a", 10)
a
```

> This is a recursive function that attempts to re-bind the value to the variable in the current environment, followed by the parent, followed by it's parent, etc. It re-binds in the first environment it finds where the variable exists. This will recurse all the way past the `.GlobalEnv` and through the package environments, unlike `<<-` which stops at the `.GlobalEnv` and creates the variable if it doesn't exist there.


