---
title: "Foundations - Names and Values"
output: rmarkdown::github_document
---

# Subsetting #

## 4.2 Selecting multiple elements  ##
1. Fix each of the following common data frame subsetting errors:

```{r 4.2.1, eval=FALSE}
mtcars[mtcars$cyl = 4, ]
mtcars[-1:4, ]
mtcars[mtcars$cyl <= 5]
mtcars[mtcars$cyl == 4 | 6, ]
```

2. Why does the following code yield five missing values? (Hint: why is it different from x[NA_real_]?)

```{r 4.2.2, eval=FALSE}
x <- 1:5
x[NA]
#> [1] NA NA NA NA NA
```
3. What does upper.tri() return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?

```{r 4.2.3, eval=FALSE}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```

4. Why does mtcars[1:20] return an error? How does it differ from the similar mtcars[1:20, ]?

> A: t does not know if 1:20 are the rows or columns. 

5.Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).

> A:

6. What does df[is.na(df)] <- 0 do? How does it work?

> A: Replaces all NAs with 0s


## 4.3 Selecting a single element  ##

1. Brainstorm as many ways as possible to extract the third value from the cyl variable in the mtcars dataset.

> A:

2. Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Then extract the R squared from the model summary (summary(mod)).

> A:

## 4.5 Applications  ##



1. Positive integers select elements at specific positions, negative integers drop elements; logical vectors keep elements at positions corresponding to TRUE; character vectors select elements with matching names.

> A:

2. [ selects sub-lists: it always returns a list. If you use it with a single positive integer, it returns a list of length one. [[ selects an element within a list. $ is a convenient shorthand: x$y is equivalent to x[["y"]].

> A:

3. Use drop = FALSE if you are subsetting a matrix, array, or data frame and you want to preserve the original dimensions. You should almost always use it when subsetting inside a function.

> A:

4. If x is a matrix, x[] <- 0 will replace every element with 0, keeping the same number of rows and columns. In contrast, x <- 0 completely replaces the matrix with the value 0.

> A:

5. A named character vector can act as a simple lookup table: c(x = 1, y = 2, z = 3)[c("y", "z", "x")]

> A:
