---
title: "Foundations: Names and Values"
output: rmarkdown::github_document
---

- copy-on-modify modify behaviour is important

- Trace copy-on-modify behaviour: 
- `tracemem()`
- `untracemem()`

- Lists: more complex because instead of storing the values itself, it instead stores references to them

- "This, however, is a shallow copy: the list object and its bindings are copied, but the values pointed to by the bindings are not."

- To see values that are shared across lists, use `lobstr::ref()`

- Data frames are lists of vectors

- Character vectors are strange! "But this is a polite fiction. R actually uses a global string pool where each element of a character vector is a pointer to a unique string in the pool:"

- find out how much memory object takes with `lobstr::obj_size()`

- "because R uses a global string pool character vectors take up less memory than you might expect"

- "obj_size(x) + obj_size(y) will only equal obj_size(x, y) if there are no shared values."

- Modify-in-place: "As we’ve seen above, modifying an R object usually creates a copy. There are two exceptions: 
Objects with a single binding get a special performance optimisation.
Environments, a special type of object, are always modified in place"

- hard to predict whether or not a copy will occur; base R functions create exceptions

- In comparison to using data frames in a loop, "We can reduce the number of copies by using a list instead of a data frame. Modifying a list uses internal C code, so the refs are not incremented and only a single copy is made:"

Lots of copies:

```{r}
x <- data.frame(matrix(runif(5 * 1e4), ncol = 5))
medians <- vapply(x, median, numeric(1))

for (i in seq_along(medians)) {
  x[[i]] <- x[[i]] - medians[[i]]
}
```

Only a single copy:

```{r}
y <- as.list(x)
cat(tracemem(y), "\n")
#> <0x7f80c5c3de20>
  
for (i in 1:5) {
  y[[i]] <- y[[i]] - medians[[i]]
}
```

- environments are always modified in place

- lobstr::mem_used() is a wrapper around gc() that prints the total number of bytes used
